---
title: "GO_and_KEGG_enrichment_analysis_Purkinje_cells"
author: "Dario"
format: html
editor: visual
---

## Note

This script uses 324 Purkinje cell-enriched genes derived from pseudobulking analysis of Stereo-seq data. How the selection was performed can be found in the script: pseudobulk_analysis_plots.py.

## GO term and KEGG enrichment analysis of 342 Purkinje cell-enriched genes

-   load libraries

```{r}
library(org.Hs.eg.db) # to get human ensembl gene annotations
library(jsonlite) # for importing json file
library(clusterProfiler) # for enrichment analyses
library(DOSE) # for visualizing enrichment results
```

-   load files (only works when working in the right directory)

```{r}
setwd("/Users/dario/Desktop/")

purkinje_genes <- fromJSON("Purkinje_genes.json")
cerebellum_genes <- fromJSON("Cerebellum_genes.json")
length(purkinje_genes)
length(cerebellum_genes)
```

### GO term enrichment analysis (overrepresentation analysis)

```{r}
ora_results <- enrichGO(gene = purkinje_genes, universe = cerebellum_genes,
                        OrgDb = org.Hs.eg.db, keyType = "SYMBOL", ont = "ALL",          
                        pAdjustMethod = "BH", pvalueCutoff  = 0.05, qvalueCutoff  = 0.2)

print(head(ora_results))
```

```{r echo=TRUE, fig.width=6.5, fig.height= 6}
dotplot(ora_results, showCategory = 10, title = "Enriched GO Terms")
```

### KEGG pathway enrichment analysis

```{r}
# convert gene names to ENTREZ ids
cerebellum_genes_entrez <- bitr(cerebellum_genes, fromType = "SYMBOL", toType = "ENTREZID",
                                OrgDb = org.Hs.eg.db)
purkinje_genes_entrez <- bitr(purkinje_genes, fromType = "SYMBOL", toType = "ENTREZID",
                              OrgDb = org.Hs.eg.db)

# perform enrichment
kegg_results <- enrichKEGG(gene = purkinje_genes_entrez$ENTREZID, universe = cerebellum_genes_entrez$ENTREZID, 
                           organism = 'hsa', pvalueCutoff  = 0.05, qvalueCutoff  = 0.2)

head(as.data.frame(kegg_results))
```

```{r echo=TRUE, fig.width=6.5, fig.height= 6}
dotplot(kegg_results, showCategory = 10, title = "Enriched KEGG Pathways")
```
