---
title: "Upset_plot_enriched_purkinje_genes"
author: "Dario"
format: html
editor: visual
---

## Note:

This script is based on prior analysis of matching Purkinje cell-enriched genes with Human Protein Atlas data, and uses matches (across subcellular locations) to perform a UpSetPlot. How the gene sets were derived can be found in the script: match_stereo_with_hpa_manual_annotations.py

## UpSet Plot for comparing overlaps in protein subcellular localization in Purkinje cells

-   load necessary library

```{r}
library(UpSetR)
```

-   set working directory and import gene lists

```{r}
setwd("/Users/dario/Desktop/purkinje_project/Task_2/Cell_segmentation_scripts")
genes_enr_in_cyto <- scan("genes_enr_in_cyto.txt", what = "", sep = "\n")
genes_enr_in_dendrites <- scan("genes_enr_in_dendrites.txt", what = "", sep = "\n")
genes_enr_in_nucleus <- scan("genes_enr_in_nucleus.txt", what = "", sep = "\n")
```

-   create input list aligning with necessary format for creating the UpSet plot

```{r}
listInput <- list("High/Medium Expression - Cytoplasm" = genes_enr_in_cyto,
                  "High/Medium Expression - Dendrites" = genes_enr_in_dendrites,
                  "High/Medium Expression - Nucleus" = genes_enr_in_nucleus)
```

-   create UpSetPlot

```{r, fig.height = 6, fig.width = 12}
upset(fromList(listInput), 
      order.by = "freq", 
      text.scale = c(1, 1, 1, 1, 1, 1)*2)
```
